Geometric variational problems

### 2020-11-14 NSF2020/Description/preliminarywork.tex

\section{Preliminary work}
\label{sec:preliminary_work}
The HAP project was initiated in 2017 by PIs RR and YNY.
The lead author on our first paper (\cite{Fu2018_SIAM}) was Szu-Pei Fu,
who was PI YNS's PhD student and is currently a postdoc with PI RR.
The collaboration involved theoretical work and simulation. The theoretical
work included model development, proving the first variation formula
\eqref{stress} and force-free law, and an energy-based  uniqueness
theorem for the exterior domain problem \eqref{SL}.
Coauthors Andreas Kl\"ockner (Computer Science, UIUC, NSF CAREER \#1654756)
and his PhD student Matt Wala provided their Quadrature by Expansion, or QBX
algorithm and Fu ran simulations on NJIT clusters.

The simulation results were encouraging. The small particle-number
simulations successfully demonstrated gradient-driven self-assembly of amphiphilic
particles into two-dimensional micelles, bilayer membranes. The model parameters
where the decay length $\rho$, interfacial tension $\gamma$, particle geometry, and
repulsion strength. The particle diamaters were set equal to monolayer thickness, 
2.0 nm \cite{Boal}, and we chose a repulsion strength that gave adequate interparticle 
spacing. For the decay length, we used $\rho = 2.5$ nm based on measurements of hydrophobic
attraction between surfactant-coated \cite{Eriksson1989,Lin2005,Parsegian,Israelachvili80}.
For the interfacial tension, we used $\gamma = 4$ pN nm$^{-1}$
based on measurements for single-component bilayer lipid membranes
\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}.

\begin{wraptable}[10]{r}{0.55\textwidth}
\begin{center}
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & exper. value          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
\end{center}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
We performed three types of simulations. The first measured bending by loading
a planar bilayer. The second used a harmonic bond to dilate a circular bilayer.
The third measured tilt using a decay equation from  \cite{KUZMIN2005}.
These simulations isolated the deformations of \eqref{ansatz3}, enabling
us to read off elastic moduli from simulation data. The results agreed remarkably 
well with the values reported in the experimental literature, 
Table~\ref{tab:moduli}. The agreement is underscored by the fact
that the two, main HAP parameters $\gamma$ and $\rho$ were assigned
physical values from the outset, rather than being tuned.

The PIs BQ, RR and YNY initiated a collaboration in summer, 2020 to develop
the hydrodynamic model of \S \ref{subsec:specific_aim_3}
To date, we have written solvers for the two-dimensional boundary integral formulatons
of \eqref{SL} and \eqref{eq:stokes}. The solvers gives third-order spatial accuracy
up to the boundary for $u$ and $\mathbf{u}$. The specialized reciprocal identity \ref{eq:reciprocal}
gives spectral accuracy for force and torque (Table \ref{tab:spectral_force}}).
\begin{wraptable}[10]{l}{0.45\textwidth}
\begin{center}
  \begin{tabular}{|l|l|l|l|}
 $n$   & $\max_{i} \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|$   &  $\max_{i} \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|$   \\
 32   &  81.4933     &   70.3970\\       
  64  &   4.9328     &    3.1928 \\
 128  &   0.0125     &    0.0030 \\
256   &   0.0000765   &   0.00003389        
  \end{tabular}
\end{center}
\caption{\label{tab:spectral_force} 
\footnotesize 
Numerical error for force and torque as a function of number of grid points $n$ per particle.
The data is for $N = 5$ particles characteristic diamater 2; two of the particles are nearly touching 
at a distance 0.02.} 
\end{wraptable}

We have simulated two-dimensional vesicles in shear flow 
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate $\dot{gamma}$. 
The particle-based system mimiks the behavior observed in continuous vesicles
\cite{torres-sanchez_millan_arroyo_2019, mahapatra_saintillan_rangamani_2020, 
Steigmann99, C6SM02452A}. The basic pattern is that at moderate shear rates,
the initially circular vesicle bilayer elongates and approaches a steadily 
tank-treading ellipse. Analyzing any underlying constituative laws, the particles
reorganize so that there is an initial decrease in vesicle area that plateaus 
to a steady state area, respectively increase in arc length. This suggests that
the particle system behaves as a semipermeable vesicle \cite{323e9a2f0c58487ea82518d7a1f96485,YAO2017728}.
Our preliminary results already showed that the particle-based bilayers have the same
large area modulus as real lipid bilayer membranes, and permit a small amount of stretching. 
For large shear rates, the vesicle ruptures and flow carries away segmented membrane patches.

The above considerations justify using the particle-based model to study material properties 
as outline in \S \ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}.
Intermonolayer slip enters zero-thickness membrane models as a velocity
jump boundary condition \cite{schwalbe_vlahovska_miksis_2010}, for example, 
but in the present model is a consequence of monolayer independence.

>>>>

\begin{wraptable}[8]{r}{0.55\textwidth}
\centering{
  \begin{tabular}{|l|r|r|p{0.1\textwidth}|}
    \hline
          & HAP                     & experiment          & ref.\\
    \hline 
    $\KB$ & $11 \pm 2.5$ \kBT       & 10 \kBT               & \cite{Naetal15,VeBrPa15,NAGLE2000159,PhysRevLett.113.248102}\\
    \hline 
    $\KA$ & $34$ \kBT \; nm$^{-2}$  & 30--40 \kBT nm$^{-2}$ & \cite{Nagle17, Nagle17-2}\\
    \hline 
    $\KTH$ & $12$ \kBT \; nm$^{-2}$ & 10 \kBT nm$^{-2}$     & \cite{KUZMIN2005, KoNa15} \\ \hline
  \end{tabular}
}
  \vspace{-8pt}
\caption{\label{tab:moduli} \footnotesize Comparision of values elastic
  moduli from the experimental literature and values derived by HAP
  simulation.} 
\end{wraptable}
\section{Preliminary work}
\label{sec:preliminary_work}
The HAP project was initiated in 2017 by PIs RR and YNY. The lead author
on our first paper~\cite{Fu2018_SIAM} was Szu-Pei Fu, who was PI YNY's
PhD student and is currently a postdoc with PI RR. The collaboration
involved theoretical work and simulation. The theoretical work included
model development, proving the first variation formula~\eqref{stress}
and force-free law, and an energy-based uniqueness theorem for the
exterior domain problem~\eqref{SL}. Coauthors Andreas Kl\"ockner
(Computer Science, UIUC, NSF CAREER \#1654756) and his PhD student Matt
Wala provided their Quadrature by Expansion, or QBX algorithm and Fu ran
simulations on NJIT clusters.

\begin{wraptable}[11]{l}{0.42\textwidth}
\centering{
  \begin{tabular}{|l|l|l|l|}
    $n$   & (rel.~err.)$_F$
    &  (rel.~err.)$_G$ \\
    \hline
% 32   &   4.07466     &   1.56437\\
% 64   &   0.24664     &   0.03476\\
% 128  &   0.00063     &   0.00006\\
  32  & $4.07 \times 10^{+0}$ & $1.56 \times 10^{+0}$ \\
  64  & $2.47 \times 10^{-1}$ & $3.48 \times 10^{-2}$ \\
  128 & $6.30 \times 10^{-4}$ & $6.00 \times 10^{-5}$ 
  \end{tabular}}
\vspace{-5pt}
\caption{\label{tab:spectral_force} 
\footnotesize  Relative numerical errors (rel.~err.)$_F = \max_{i}
  \|\mathbf{F}_i-\mathbf{F}_i^{\text{exact}}\|/\|\mathbf{F}_i^{\text{exact}}\|$
  and (rel.~err.)$_G = \max_{i}
  \|\mathbf{G}_i-\mathbf{G}_i^{\text{exact}}\|/\|\mathbf{G}_i^{\text{exact}}\|$
  for force and torque respectively as a function of number of grid
  points $n$ per particle.  The data is for $N = 5$ particles; two of
  the particles are nearly touching at a distance 1\% of particle
  diameter.} 
\end{wraptable}
The simulation results were encouraging. Small particle-number
simulations successfully demonstrated gradient-driven self-assembly of
amphiphilic particles into two-dimensional micelles and bilayer membranes.
The model parameters were the decay length $\rho$, interfacial tension
$\gamma$, particle geometry, and repulsion strength. The particle
diameters were the physical monolayer thickness, 2.0 nm~\cite{Boal}, and
we chose a repulsion strength that gave adequate interparticle spacing.
For the decay length, we used $\rho = 2.5$ nm based on measurements of
hydrophobic attraction between surfactant-coated
membranes~\cite{Eriksson1989, Lin2005, Parsegian, Israelachvili80}. For
the interfacial tension, we used $\gamma = 4$ pN nm$^{-1}$ based on
measurements for single-component bilayer lipid
membranes~\cite{GarciaSaez, KUZMIN2005, Petelska2012, Jackson2016}.

\begin{wrapfigure}[]{r}{0.5\textwidth}
  \centering{\includegraphics[width=0.5\textwidth]{Figures/TankTreading.pdf}}
  \vspace{-25pt}
    \caption{\label{fig:tank_treading} \footnotesize A two-dimensional
    particle-based vesicle in a shear flow. Panel A shows the particles,
    water activity, and background flow direction. Panel B plots the
    mean angular velocities for the outer leaflet and the inner
    respectively.  Panel C shows that the trajectory of the particle
    centers forms two ellipses with well-defined inclination angle
    $\Theta$. Panel D shows the evolution of the inclination angle.
    Panels B and D share the same time axis.}
\end{wrapfigure}
The PIs BQ, RR and YNY initiated a collaboration in Summer 2020 to
develop the hydrodynamic model of \S\ref{subsec:specific_aim_3}.  To
date, we have written solvers for the two-dimensional boundary integral
formulations of~\eqref{SL} and~\eqref{eq:stokes} that achieve
third-order accuracy up to the boundary. The specialized reciprocal
identity~\eqref{eq:reciprocal} gives spectral accuracy for force and
torque (Table~\ref{tab:spectral_force}).

We performed three types of simulations. The first measured bending by
loading a partially clamped planar bilayer. The second used a harmonic
bond to dilate a circular bilayer. The third measured tilt using a decay
equation from~\cite{KUZMIN2005}. These simulations isolated three of the
five deformations of~\eqref{ansatz3}, enabling us to read off elastic
moduli from simulation data. The results agreed remarkably well with the
values reported in the experimental literature (Table~\ref{tab:moduli}).
The agreement is underscored by the fact that the two main HAP
parameters, $\gamma$ and $\rho$, were assigned physical values from the
outset rather than being tuned to fit data.

We have simulated two-dimensional vesicles in shear flow
$\mathbf{u}_{\infty} = \dot{\gamma} x_2 \mathbf{i}_1$ with shear rate
$\dot{\gamma}$,
%(The symbol shear rate is unrelated to interfacial tension.)
and the particle-based system mimics the behavior observed in continuous
vesicles~\cite{torres-sanchez_millan_arroyo_2019,
mahapatra_saintillan_rangamani_2020, Steigmann99, C6SM02452A}. The basic
pattern is that the vesicle bilayer approaches a steady tank-treading
ellipse (Figure~\ref{fig:tank_treading}C). We analyzed the simulation
data in the context of theoretical work on
tank-treading~\cite{Finken2008}. Jeffrey's equation predicts a
tank-treading frequency $\dot{\gamma}/2$ (dashed line in
Figure~\ref{fig:tank_treading}B). In the particle simulation, the
tank-treading frequency corresponds to the mean of the particle angular
velocities, and these approach a value within 7\% of the theoretical
prediction (dotted and solid curves in Figure~\ref{fig:tank_treading}B).
In the low viscosity-contrast regime of our simulation, Jeffrey's
equation predicts a decrease in the ellipse's inclination angle $\Theta$
and our particle-based simulation follows this prediction
(Figure~\ref{fig:tank_treading}D). The higher frequency oscillation
comes from particle reorganization.

Analyzing for any underlying constitutive laws, the particles reorganize
so that there is an initial increase and then plateauing of vesicle
circumference. As is the case with real lipid bilayer membranes, the
particle collection permits a small amount of stretching. The
theoretical calculations give an increase of length by 3.6\% which
agrees with our measured increase in length. The simulations also
suggest that the particle system behaves as a semipermeable
vesicle~\cite{323e9a2f0c58487ea82518d7a1f96485, YAO2017728}.

\begin{wrapfigure}[11]{r}{0.22\textwidth}
  \centering{\includegraphics[width=0.22\textwidth]{Figures/PW_fig5.pdf}}
  \caption{\label{fig:flow_rupture} \footnotesize Rupture of a
  two-dimensional vesicle at large shear rates.}
\end{wrapfigure}
The above considerations further justify using the particle-based model
to study material properties as outlined in
\S\ref{subsec:specific_aim_1}. Furthermore, the particle system allows
for intermonolayer slip~\cite{SHKULIPA2005823, ShkulipaThesis}. In
Figure \ref{fig:tank_treading}B, the inner angular velocity is slightly
less than the outer angular velocity. Intermonolayer slip enters
zero-thickness membrane models as a velocity jump boundary condition
\cite{schwalbe_vlahovska_miksis_2010}, for example, but in the present
model, it is a consequence of monolayer independence. For large shear
rates, viscous forces exceed the hydrophobic attraction and the vesicle
ruptures; the flow carries away segmented membrane patches (Figure
\ref{fig:flow_rupture}).
